---
title: "statistical-analysis"
format: html
editor: visual
---

## 

```{r}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
#install.packages("maps")
library(maps)
#install.packages("ggcorrplot")
library(ggcorrplot)
#install.packages("plm")
library(plm)
#install.packages("gt")
library(gt)
#install.packages("car")
library(car)
#install.packages("MASS") 
library(MASS)
#install.packages("tidyverse")
library(tidyverse)
#install.packages("reshape2")  
library(reshape2) 
#install.packages("scales")
library(scales)
library(splines)
library(here)
library(knitr)
#install.packages("kableExtra")
library(kableExtra)
#install.packages("GGally")
library(GGally)
#install.packages("patchwork")
library(patchwork)
#install.packages("gridExtra") 
library(gridExtra)
#install.packages("lmerTest")
library(lmerTest)
#install.packages("stargazer")
library(stargazer)
#install.packages("lmtest")
library(lmtest)
#install.packages("mgcv")
library(mgcv)
#install.packages("Metrics")
library(Metrics)
#install.packages("reticulate")
library(reticulate)
#install.packages("xgboost")
library(xgboost)
#install.packages("tidymodels")y
library(tidymodels)
library(Metrics)
library(gamm4)
library(broom)
#install.packages("gratia")
library(gratia)
library(dplyr)


```

Load the dataset

```{r}

asia_africa_location <- here("data", "processed-data", "seaa.rds")
asia_africa <- readRDS(asia_africa_location)

asia_africa <- asia_africa %>%
  filter(country != "Lesotho") #drop Lesotho, it is potential influential outlier 


colnames(asia_africa)


asia_africa$pro_basicwater <- as.numeric(asia_africa$pro_basicwater)
asia_africa$pro_safewater <- as.numeric(asia_africa$pro_safewater)
asia_africa$pro_basicsanitation <- as.numeric(asia_africa$pro_basicsanitation)
asia_africa$pro_safesanitation <- as.numeric(asia_africa$pro_safesanitation)
asia_africa$country <- as.factor(asia_africa$country)
str(asia_africa)
```

## Data Analysis

```{r}
# Adding Fixed-Effect to Make Conditional Models ---
ggplot(asia_africa, aes(x = year, y = tbi)) + 
  geom_line(aes(group=country), col="grey") + 
  stat_smooth(method="loess", col="black", lwd=1.5, se=FALSE)+
  scale_x_continuous(name = "years") + 
  scale_y_continuous(name = "Tuberculosis Incidence") +
  theme_bw() + 
  theme(axis.text=element_text(size=14, colour="black"), 
        axis.title=element_text(size=14,face="bold")) +
  theme(strip.text.x = element_text(size = 14))+ theme(legend.position="none") 
```

```{r}
# Create scatter plot faceted by independent variable
scatter_faceted <- ggplot(asia_africa_long, aes(x = value, y = tbi, color = country)) + 
  geom_point(alpha = 0.6) +  # Scatter points, colored by country
  stat_smooth(method = "loess", col = "black", lwd = 1.2, se = FALSE) +  # LOESS (non-linear) in black
  facet_wrap(~ variable, scales = "free_x") +  # Facet by variable
  labs(
    title = "TBI and Its Association with Water, Sanitation, and HDI Indicators",
    x = "Independent Variable Value",
    y = "Tuberculosis Incidence (TBI)"
  ) +
  theme_bw() + 
  theme(axis.text = element_text(size = 12, colour = "black"), 
        axis.title = element_text(size = 14, face = "bold"),
        strip.text = element_text(size = 14, face = "bold"), 
        legend.position = "none")  # Hides the legend


#save the plot 
plots_tbi_vs_allvars_loc <- here( "results", "figures", "analysis", "plots_tbi_vs_allvars.png") # to set up location for the pictures created 
ggsave(filename = plots_tbi_vs_allvars_loc  , plot=scatter_faceted, width = 16, height = 12, units = "in", dpi = 300) # save the pictures created 

print(scatter_faceted)

```

Non-linearity test

```{r}
# Compute the correlation matrix for selected variables
cor(asia_africa[, c("tbi", "pro_basicwater", "pro_safewater", 
                    "pro_basicsanitation", "hdi")])

# Fit a linear model
lm_model <- lm(tbi ~  pro_basicwater + pro_safewater + 
               pro_basicsanitation + hdi, 
               data = asia_africa)

# Run the RESET test for linearity
resettest(lm_model, power = 2:3)

```

```{r}
# Fit a linear model (LM)
lm_model <- lm(tbi ~ pro_basicwater + pro_safewater + 
               pro_basicsanitation +  hdi, 
               data = asia_africa)

# Fit a non-linear model (GAM)
gam_model <- gam(tbi ~ s(pro_basicwater) + s(pro_safewater) + 
                  s(pro_basicsanitation)  + s(hdi), 
                 data = asia_africa, method = "REML")

# Compare AIC values
AIC(lm_model, gam_model)


# Plot residuals vs. fitted values for the linear model
plot(lm_model$fitted.values, resid(lm_model), 
     xlab = "Fitted Values", ylab = "Residuals", 
     main = "Residuals vs. Fitted Plot")
abline(h = 0, col = "red")  # Horizontal line at 0

```

```{r}

gamm_model_interactions<- gam(
  tbi ~ 
    s(year) + 
    s(pro_basicwater) + 
    s(pro_safewater) + 
    s(pro_basicsanitation) + 
    s(pro_safesanitation) + 
    s(hdi) +  
    ti(pro_basicwater, hdi) + 
    ti(pro_safewater, hdi) +
    ti(pro_basicsanitation, hdi) + 
    ti(pro_safesanitation, hdi) +
    ti(pro_safesanitation, pro_basicwater) +
    ti(pro_safewater, pro_basicwater) +
    s(country, bs = "re"),         
  family = nb(),                  
  method = "REML",                 
  data = asia_africa
)


summary(gamm_model_interactions)
AIC(gamm_model_interactions)

```

Note:

The warning indicate that the model is too complex, I will delete the least significant interactions:

```{r}

gamm_model_interactions_reduced<- gam(
  tbi ~ 
    s(year) + 
    s(pro_basicwater) + 
    s(pro_safewater) + 
    s(pro_basicsanitation) + 
    s(pro_safesanitation) + 
    s(hdi) +  
    ti(pro_basicwater, hdi) + 
    ti(pro_safewater, hdi) +
    ti(pro_safesanitation, pro_basicwater) +
    ti(pro_safewater, pro_basicwater) +
    s(country, bs = "re"),         
  family = nb(),                  
  method = "REML",                 
  data = asia_africa
)


AIC(gamm_model_interactions_reduced)

summary(gamm_model_interactions_reduced)

```

### Visualize the result of the final model

```{r}
# Extract summary of the model
model_summary <- summary(gamm_model_interactions_reduced)

# Parametric terms
param_tbl <- as.data.frame(model_summary$p.table) %>%
  tibble::rownames_to_column("term") %>%
  rename(
    estimate = Estimate,
    std.error = `Std. Error`,
    z.value = `z value`,
    p.value = `Pr(>|z|)`
  ) %>%
  mutate(
    type = "Parametric",
    edf = NA, 
    chi_sq = NA
  )

# Smooth terms
smooth_tbl <- as.data.frame(model_summary$s.table) %>%
  tibble::rownames_to_column("term") %>%
  rename(
    edf = edf,
    chi_sq = `Chi.sq`,
    p.value = `p-value`
  ) %>%
  mutate(
    type = "Smooth",
    estimate = NA, 
    std.error = NA, 
    z.value = NA
  )

# Combine parametric and smooth tables
model_results_tbl <- combined_tbl %>%
  dplyr::select(type, term, estimate, std.error, edf, chi_sq, p.value)


# Display table with gt
model_results_tbl %>%
  gt(groupname_col = "type") %>%
  fmt_number(columns = where(is.numeric), decimals = 3) %>%
  tab_header(
    title = "Summary of GAMM Estimates",
    subtitle = "Parametric and Smooth Terms"
  ) %>%
  cols_label(
    term = "Term",
    estimate = "Estimate",
    std.error = "Std. Error",
    edf = "EDF",
    chi_sq = "Chi-square",
    p.value = "p-value"
  ) %>%
  sub_missing(everything(), missing_text = "-")


#save the table 

# Save as RDS
saveRDS(model_results_tbl, file = here::here("results", "tables", "analysis", "gamm_summary_table.rds"))

```

NULL Model

```{r}
model_null_gamm <- gamm(tbi ~ 1, random = list(country = ~1), data = asia_africa)
```

```{r}

AIC(lm_model)
AIC(gam_model)
AIC(gamm_model_interactions)
AIC(gamm_model_interactions_reduced)

```

Based on the table, although Generalized model with full interaction results the smallest AIC, its df is negative indicating the model instability. Therefore, we used the GAMM with reduced model, which least significant interaction is reduced.

***Visualize the AIC***

```{r}
# create AIC data frame
aic_results <- data.frame(
  Model = c("LM", "GAM", "GAMM + Interactions", "GAMM + Interactions (Reduced)"),
  AIC = c(4735.692, 4568.658, 2956.188, 2954.157)
)


# create plot 
plot_aic <- ggplot(aic_results, aes(x = reorder(Model, AIC), y = AIC)) +
  geom_point(size = 4, color = "darkgreen") +
  geom_segment(aes(xend = Model, yend = 0), linetype = "dashed") +
  geom_text(aes(label = round(AIC, 1)), hjust = -0.2, size = 4) +
  coord_flip() +
  labs(title = "AIC by Model", x = "Model", y = "AIC") +
  theme_minimal()

plot(plot_aic)
```

\*\*\* Print RMSE from those models\*\*\*

```{r}

# 1. LM Model
pred_lm <- predict(lm_model)
rmse_lm <- sqrt(mean((asia_africa$tbi - pred_lm)^2))

# 2. GAM Model
pred_gam <- predict(gam_model, type = "response")
rmse_gam <- sqrt(mean((asia_africa$tbi - pred_gam)^2))


# 4. GAMM Model with Interactions
pred_gamm_interactions <- predict(gamm_model_interactions, type = "response")
rmse_gamm_interactions <- sqrt(mean((asia_africa$tbi - pred_gamm_interactions)^2))

# 5. Reduced GAMM Model with Interactions
pred_gamm_interactions_red <- predict(gamm_model_interactions_reduced, type = "response")
rmse_gamm_interactions_red <- sqrt(mean((asia_africa$tbi - pred_gamm_interactions_red)^2))

# 6. NULL Model (GAMM null model)
pred_null <- predict(model_null_gamm$gam, type = "response")
rmse_null <- sqrt(mean((asia_africa$tbi - pred_null)^2))



# Print results
rmse_results <- data.frame(
  Model = c("Null Model", "LM", "GAM",  "GAMM + Interactions", "GAMM + Interactions (Reduced)" ),
  RMSE = c(rmse_null, rmse_lm, rmse_gam, rmse_gamm_interactions, rmse_gamm_interactions_red )
)

saveRDS(rmse_results, file = here("results", "tables", "analysis", "rmse_results.rds"))

print(rmse_results)
```

Althought GAMM with full interaction has smallest RMSE, based on the negative value of DF. we chose Gamm with interactons (reduced)

```{r}

#| label: tbl-rmse
#| tbl-cap: "RMSE Comparison of Models"
#| echo: false

# Load the table
rmse_results <- readRDS(here("results", "tables", "analysis", "rmse_results.rds"))

# Display the table
kable(rmse_results, digits = 2)
```

Vizualise the RMSE

```{r}


plot_RMSE <- ggplot(rmse_results, aes(x = reorder(Model, RMSE), y = RMSE)) +
  geom_point(size = 4, color = "darkred") +
  geom_segment(aes(xend = Model, yend = 0), linetype = "dashed") +
  geom_text(aes(label = round(RMSE, 2)), 
            hjust = -0.2, size = 4, color = "black") +
  coord_flip() +
  labs(title = "RMSE by Model", x = "Model", y = "RMSE") +
  theme_minimal()

ggsave(
  filename = here("results", "figures", "analysis", "rmse_by_model.png"),
  plot = plot_RMSE,
  width = 7, height = 5, dpi = 300
)

plot(plot_RMSE)

# facet the AIC and RMSE plots 
Plot_AIC_RMSE <- plot_aic + plot_RMSE + plot_layout(ncol = 2)




# Save the plot 
ggsave(
  here("results", "figures", "analysis", "plot_aic_rmse.png"),
  plot = Plot_AIC_RMSE,
  width = 12, height = 8, dpi = 300
)

plot(Plot_AIC_RMSE)
```

Predicted and observed TBI from original dat

```{r}

asia_africa <- asia_africa %>%
  mutate(predicted_tbi = predict(gamm_model_interactions_reduced, newdata = asia_africa, type = "response"))


#Create plot
plot_observed_predicted_original <- ggplot(asia_africa, aes(x = predicted_tbi, y = tbi)) +
  geom_point(alpha = 0.6, color = "#1f77b4", size = 2) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray30") +
  labs(
    title = "Predicted vs Observed TBI (original data)",
    x = "Predicted TBI",
    y = "Observed TBI"
  ) +
  theme_minimal(base_size = 14)

# Save the plot 
ggsave(
  here("results", "figures", "analysis", "plot_observed_predicted_original.png"),
  plot = plot_observed_predicted_original,
  width = 10, height = 8, dpi = 300
)

plot(plot_observed_predicted_original)

```

#Machine Learning read the dataset

*Splitted Dataset*

```{r}
# create setseed for reproductibility 
ml_seed <- set.seed(2345)

# Split dataset into a 75% train set and a 25% test set

splitted_ml <- initial_split(asia_africa, prop= .75)
df_train <- training(splitted_ml)
df_test <- testing(splitted_ml)

```

***fit the models in to machine learning***

The study is supposed to use Mixed Effect Random Forest (MERF). Since there is no a specific package for MERF in R,I will use xgboost

***Linear Model***

```{r}
# Fit a linear model (LM) with train data
lm_model_train <- lm(tbi ~ pro_basicwater + pro_safewater + 
               pro_basicsanitation + pro_safesanitation +  hdi, 
               data = df_train)

# Fit a non-linear model (GAM)
gam_model_train <- gam(tbi ~ s(pro_basicwater) + s(pro_safewater) + 
                  s(pro_basicsanitation) + s(pro_safesanitation) + s(hdi), 
                 data = df_train, method = "REML")

# Fit Gamm Model without interaction with training data 
gamm_model_train <- gam(
  tbi ~ 
    s(year) + 
    s(pro_basicwater) + 
    s(pro_safewater) + 
    s(pro_basicsanitation) + 
    s(pro_safesanitation) + 
    s(hdi) +  
    s(country, bs = "re"),         
  family = nb(),                  
  method = "REML",                 
  data = df_train
)

summary(gamm_model_train)

# Fit GamM Model with interaction  with training data 


gamm_model_interactions_train <- gam(
  tbi ~ 
    s(year) + 
    s(pro_basicwater) + 
    s(pro_safewater) + 
    s(pro_basicsanitation) + 
    s(pro_safesanitation) + 
    s(hdi) +  
    ti(pro_basicwater, hdi) + 
    ti(pro_safewater, hdi) +
    ti(pro_basicsanitation, hdi) + 
    ti(pro_safesanitation, hdi) +
    ti(pro_safesanitation, pro_basicwater) +
    ti(pro_safewater, pro_basicwater) +
    s(country, bs = "re"),         
  family = nb(),                  
  method = "REML",                 
  data = df_train
)

summary(gamm_model_interactions_train)

```

***Fit generated data with final model***

```{r}
# Fit GamM Model with interaction (reduced) with training data 
gamm_model_interactions_reduced_train  <- gam(
  tbi ~ 
    s(year) + 
    s(pro_basicwater) + 
    s(pro_safewater) + 
    s(pro_basicsanitation) + 
    s(pro_safesanitation) + 
    s(hdi) +  
    ti(pro_basicwater, hdi) + 
    ti(pro_safewater, hdi) +
    ti(pro_safesanitation, pro_basicwater) +
    ti(pro_safewater, pro_basicwater) +
    s(country, bs = "re"),         
  family = nb(),                  
  method = "REML",                 
  data = df_train
)

summary(gamm_model_interactions_reduced_train  )

```

**Fit Xgboost with train data**\*

```{r}

# Manually create interaction terms (to mimic GAMM)
df_train <- df_train %>%
  mutate(
    int_basicwater_hdi        = pro_basicwater * hdi,
    int_safewater_hdi         = pro_safewater * hdi,
    int_safesanitation_basic  = pro_safesanitation * pro_basicwater,
    int_safewater_basic       = pro_safewater * pro_basicwater
  )

df_test<- df_test %>%
  mutate(
    int_basicwater_hdi        = pro_basicwater * hdi,
    int_safewater_hdi         = pro_safewater * hdi,
    int_safesanitation_basic  = pro_safesanitation * pro_basicwater,
    int_safewater_basic       = pro_safewater * pro_basicwater
  )

# One-hot encode 'country' and create full model matrix (train data) 
X_train <- model.matrix(
  tbi ~ year + pro_basicwater + pro_safewater + pro_basicsanitation +
    pro_safesanitation + hdi + country + 
    int_basicwater_hdi + int_safewater_hdi +
    int_safesanitation_basic + int_safewater_basic,
  data = df_train
)[, -1]  # remove intercept


# Test Data 
X_test <- model.matrix(
  tbi ~ year + pro_basicwater + pro_safewater + pro_basicsanitation +
    pro_safesanitation + hdi + country + 
    int_basicwater_hdi + int_safewater_hdi +
    int_safesanitation_basic + int_safewater_basic,
  data = df_test
)[, -1]

# Extract outcome
y_train <- df_train$tbi
y_test <- df_test$tbi

# Fit XGBoost with count data (train data) 
xgb_model_train <- xgboost(
  data = X_train,
  label = y_train,
  objective = "count:poisson",
  nrounds = 100,
  eta = 0.1,
  max_depth = 6,
  subsample = 0.8,
  colsample_bytree = 0.8,
  verbose = 0
)

# Fit XGBoost with count data (testdata) 
xgb_model_test <- xgboost(
  data = X_test,
  label = y_test,
  objective = "count:poisson",  # appropriate for count outcomes
  nrounds = 100,
  eta = 0.1,
  max_depth = 6,
  subsample = 0.8,
  colsample_bytree = 0.8,
  verbose = 0
)

# Predict and evaluate


preds_xgb_train <- predict(xgb_model_train, newdata = X_train)
preds_xgb_test <- predict(xgb_model_test, newdata = X_test)


df_rmse_train <- tibble(truth = y_train, estimate = preds_xgb_train)
df_rmse_test  <- tibble(truth = y_test, estimate = preds_xgb_test)

#RMSE 
rmse(df_rmse_train, truth = truth, estimate = estimate)
rmse(df_rmse_test, truth = truth, estimate = estimate)


```

Create a plot

```{r}
# Combine training data
df_train_plot <- tibble(
  actual = y_train,
  predicted = preds_xgb_train,
  dataset = "Train"
)

plot(df_train_plot)

# Combine test data
df_test_plot <- tibble(
  actual = y_test,
  predicted = preds_xgb_test,
  dataset = "Test"
)
plot(df_test_plot)

# Combine both for plotting
df_plot <- bind_rows(df_train_plot, df_test_plot)

# Plot predicted vs observed
plot_obs_pred_ml <- ggplot(df_plot, aes(x = actual, y = predicted, color = dataset)) +
  geom_point(alpha = 0.7, size = 2) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
  facet_wrap(~ dataset) +
  labs(
    title = "Predicted vs Observed TBI",
    x = "Observed TBI",
    y = "Predicted TBI"
  ) +
  theme_minimal() +
  theme(legend.position = "none")


# Save the plot 
ggsave(
  here("results", "figures", "analysis", "plot_obs_pred_ml.png"),
  plot = plot_obs_pred_ml,
  width = 10, height = 8, dpi = 300
)

plot(plot_obs_pred_ml)
```

***Predict and Evaluate the models using test data***

```{r}
# Predict
lm_pred_test <- predict(lm_model_train, newdata = df_test, type = "response")
gam_pred_test <- predict(gam_model_train, newdata = df_test, type = "response")
gamm_pred_test <- predict(gamm_model_train, newdata = df_test, type = "response")
gamm_inte_pred_test <- predict(gamm_model_interactions_train, newdata = df_test, type = "response")
gamm_inte_pred_reduced_test <- predict(gamm_model_interactions_reduced_train, newdata = df_test, type = "response")
xgb_pred_test <- predict(xgb_model_train, newdata = X_test)

# Evaluate RMSE
rmse_ml_lm <- rmse(tibble(truth = df_test$tbi, estimate = lm_pred_test), truth, estimate)
rmse_ml_gam <- rmse(tibble(truth = df_test$tbi, estimate = gam_pred_test), truth, estimate)
rmse_ml_gamm <- rmse(tibble(truth = df_test$tbi, estimate = gamm_pred_test), truth, estimate)
rmse_ml_gamm_inte <- rmse(tibble(truth = df_test$tbi, estimate = gamm_inte_pred_test), truth, estimate)
rmse_ml_gamm_inte_reduced <- rmse(tibble(truth = df_test$tbi, estimate = gamm_inte_pred_reduced_test), truth, estimate)
rmse_ml_xgb <- rmse(tibble(truth = df_test$tbi, estimate = xgb_pred_test), truth, estimate)



```

```{r}

# Your RMSE results — replace with your actual values if different
rmse_original <- tibble(
  Model = c("Null Model", "LM", "GAM", "GAMM", "GAMM + Interactions", "GAMM + Interactions (Reduced)"),
  RMSE = c(148.58, 125.13, 85.17, 18.29, 7.77, 8.05),
  Data = "Full Data"
)

rmse_test <- tibble(
  Model = c("Null Model", "LM", "GAM", "GAMM", "GAMM + Interactions", "GAMM + Interactions (Reduced)"),
  RMSE = c(NA, 129.34, 102.64, 24.48, 10.73, 10.54),
  Data = "Test Data"
)

# Combine both into one dataframe
rmse_combined <- bind_rows(rmse_original, rmse_test)

```

```{r}


ggplot(rmse_combined, aes(x = Model, y = RMSE, fill = Data)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.6) +
  labs(
    title = "RMSE Comparison: Full Data vs Test Data",
    y = "RMSE (Lower is Better)",
    x = "Model"
  ) +
  scale_fill_manual(values = c("Full Data" = "#4E79A7", "Test Data" = "#F28E2B")) +
  theme_minimal(base_size = 13) +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))



```

Now I want to plot the predicted value vs observed value between training data and test data

```{r}
# Let's say your best model is this
gamm_model_interactions_test <- gam(
  tbi ~ 
    s(year) + s(pro_basicwater) + s(pro_safewater) +
    s(pro_basicsanitation) + s(pro_safesanitation) + s(hdi) +
    ti(pro_basicwater, hdi) + ti(pro_safewater, hdi) +
    ti(pro_safesanitation, pro_basicwater) + ti(pro_safewater, pro_basicwater) +
    s(country, bs = "re"),
  data = df_test,
  family = nb(),
  method = "REML"
)



# Predicted values
train_data_pred <- df_train %>%
  mutate(Source = "Train", .fitted = predict(gamm_model_interactions_reduced_train, type = "response"))

test_data_pred <- df_test%>%
  mutate(Source = "Test", .fitted = predict(gamm_model_interactions_test, newdata = df_test, type = "response"))

# Combine
pred_obs_df <- bind_rows(train_data_pred, test_data_pred)


# Create Plot 

plot_observed_predicted <- ggplot(pred_obs_df, aes(x = .fitted, y = tbi, color = Source, shape = Source)) +
  geom_point(alpha = 0.6, size = 2) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray50") +
  labs(
    title = "Observed vs Predicted TBI (Training vs Test Data)",
    x = "Predicted TBI",
    y = "Observed TBI"
  ) +
  theme_minimal(base_size = 14) +
  scale_color_manual(values = c("Train" = "#1f77b4", "Test" = "#ff7f0e"))

#Save the plot
ggsave(
  here("results", "figures", "analysis", "plot_observed_predicted_train_test .png"),
  plot = plot_observed_predicted ,
  width = 10, height = 8, dpi = 300
)

plot(plot_observed_predicted )

#Facet the observed vs predictied for ML data and original data 

plot_ob_pre_ml_vs_ori <- 
plot(plot_observed_predicted_original + plot_observed_predicted + plot_layout(nrow  = 2))

# Save the plot 
ggsave(
  here("results", "figures", "analysis", "plot_ob_pre_ml_vs_ori.png"),
  plot =plot_ob_pre_ml_vs_ori,
  width = 12, height = 8, dpi = 300
)

plot(plot_ob_pre_ml_vs_ori)
```

1.  Excellent fit Both training and test points fall very close to the diagonal

That means your model accurately predicts TBI, both on data it has seen and on data it hasn’t seen

This is a sign of a well-generalizing model, not overfittin

2.  Low bias & low variance Predictions are neither consistently above nor below the true values

There’s minimal scatter, indicating good precision

3.  Consistent performance across train and test The orange (test) points closely track the blue (train) points

This tells you that your model doesn't just memorize, it learns the pattern and generalizes it

#Tunning the model

```{r}

set.seed(2345)

# Cross-validation folds (same as with LASSO)
folds <- vfold_cv(asia_africa, v = 10)

#Define tunable XGBoost model
xgb_model <- boost_tree(
  trees = 1000,
  learn_rate = tune(),
  tree_depth = tune(),
  mtry = tune(),
  loss_reduction = tune(),
  sample_size = tune()
) %>%
  set_engine("xgboost", objective = "count:poisson") %>%
  set_mode("regression")

xgb_recipe <- recipe(tbi ~ ., data = asia_africa) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors())


# Create workflow

xgb_wf <- workflow() %>%
  add_model(xgb_model) %>%
  add_recipe(xgb_recipe)

# Define Tuning Grid 
xgb_grid <- grid_space_filling(
  tree_depth(),
  learn_rate(),
  loss_reduction(),
  sample_size = sample_prop(),
  finalize(mtry(), df_train),
  size = 20
)


# tune with grid search 
xgb_tuned <- tune_grid(
  xgb_wf,
  resamples = folds,
  grid = xgb_grid,
  metrics = metric_set(rmse)
)


# Visualize 
autoplot_xgb <- autoplot(xgb_tuned) 

# Save to file using here()
ggsave(
  filename = here::here("results", "figures", "analysis", "autoplot_xgb_tuning.png"),
  plot = autoplot_xgb,
  width = 10,
  height = 8,
  dpi = 300
)
```
